Delayed effects of climate on vital rates lead to demographic divergence in Amazonian forest fragments

### 2021-08-11 review/Response to reviews.Rmd

> Lines 484-486: I don't think this study strongly support that plants do worse in fragments than CF.

##  {#R2-23}

> Line 499: This is confusing.
> If you measure size every year than you calculate growth as the change in size from one year to the next, don't you?

That's correct.
This line meant to point out that size is usually measured in yearly censuses while weather data can be more easily collected at monthly or daily timescales.
We've edited this sentence to read: "The ability to identify size-specific lagged responses may be especially complicated given size is rarely measured at the same time scale (e.g. monthly) as climate drivers."

### 2021-08-11 doc/paper.Rmd
closes #89 for reals

Over 24 million ha of the Brazilian Amazon have been cleared in the last two decades [@silvajunior2021], resulting in their extensive fragmentation [@broadbentForestFragmentationEdge2008].
Climate models predict a future of extremes for these forests---increases in the frequency and geographic extent of droughts, but also increases in the frequency and area affected by periods of unusual wetness [@duffy2015].
Our results support the hypothesis that populations in Amazonian forest fragments could be more susceptible to the effects of changing climate than those in continuous forest [@lauranceApparentEnvironmentalSynergism2014].
However, they also indicate that the demographic responses to climate change of populations in fragmented landscapes may be far more complex than previously appreciated.
Multi-factorial, multi-season experiments [*sensu* @aguirre2021; @brunaCompensatoryResponsesUnderstory2005; @markewitz2010; @westerbandEarlyLifeConditions2017], ideally manipulating multiple combinations of climatic variables [@mundim2016], are needed to determine how and why habitat-specific differences in environmental conditions interact to delay the demographic responses of plants to climatic variability.
Also needed are statistical tools that can test for synergistic effects of fragmentation and climate in vital rates, as those currently available do not allow for including interaction terms.
This also limits the ability to include size by climate interactions in a DLNM; although plant responses to both fragmentation and climatic extremes can be size-specific [@brunaDemographicEffectsHabitat2005; @schwartzFragmentationForestStructure2019].
The ability to identify size-specific lagged responses may be especially complicated given size and growth are rarely measured at the same time scale as SPEI and other putative climatic drivers.

>>>>

Over 24 million ha of the Brazilian Amazon have been cleared in the last two decades [@silvajunior2021], resulting in their extensive fragmentation [@broadbentForestFragmentationEdge2008].
Climate models predict a future of extremes for these forests---increases in the frequency and geographic extent of droughts, but also increases in the frequency and area affected by periods of unusual wetness [@duffy2015].
Our results support the hypothesis that populations in Amazonian forest fragments could be more susceptible to the effects of changing climate than those in continuous forest [@lauranceApparentEnvironmentalSynergism2014].
However, they also indicate that the demographic responses to climate change of populations in fragmented landscapes may be far more complex than previously appreciated.
Multi-factorial, multi-season experiments [*sensu* @aguirre2021; @brunaCompensatoryResponsesUnderstory2005; @markewitz2010; @westerbandEarlyLifeConditions2017], ideally manipulating multiple combinations of climatic variables [@mundim2016], are needed to determine how and why habitat-specific differences in environmental conditions interact to delay the demographic responses of plants to climatic variability.
Also needed are statistical tools that can test for synergistic effects of fragmentation and climate in vital rates, as those currently available do not allow for including interaction terms.
This also limits the ability to include size by climate interactions in a DLNM; although plant responses to both fragmentation and climatic extremes can be size-specific [@brunaDemographicEffectsHabitat2005; @schwartzFragmentationForestStructure2019].
The ability to identify size-specific lagged responses may be especially complicated given size is rarely measured at the same time scale (e.g. monthly) as climate drivers.
